On modeling for Kerr black holes: Basis learning, QNM frequencies, and spherical-spheroidal mixing coefficients

### 2018-06-30 docs/mvfits_paper/mvf.tex
updating docs

\title{Notes on modeling for Kerr black holes: Basis learning, QNM frequencies, and spherical-spheroidal mixing coefficients }

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Abstract %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
	%
	The ongoing direct detection of gravitational wave signals is aided by representative models of theoretical predictions. In particular, the process of model based detection, and subsequent comparison of signals the general relativity's predictions are aided by the modeling of information related to perturbed Kerr black holes. Here, we summarize recent methods and models for the analytically understood gravitational wave spectra (quasi-normal mode frequencies), and harmonic structure of Kerr black holes (mixing coefficients between spherical and spheroidal harmonics). Towards the construction of these models, two algorithms, GMVP and GMVR, for the automated polynomial and rational modeling of general dimensional complex scalars are presented.
	%
\end{abstract}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

From the first to second line of \eqn{eq:mvt}, we have used the definition of the exponential function (i.e. its series expansion form).

@@

The adjustments of perspective result in

@@

\par The key to robustly solving for $\vec{\alpha}$ lies in the iterative refinement of $\vec{\alpha}$.

@@

\par We briefly review the application of \gmvr{} to a toy problem.

@@

We then present the application of \gmvp{} and \gmvr{} to the respective modeling of \qnm{} frequencies, and harmonic mixing coefficients.

@@

While each case is an instance of modeling on a 1D domain, we note that higher dimensional cases have been treated elsewhere \cite{London:2018gaq}.

@@

Here

@@

\subsection{Modeling \qnm{} frequencies using \gmvp{}}
In seeking to apply \gmvp{} to select \qnm{} frequencies, we wish to impose a zero-damping constraint, namely that some frequencies are real as $\jf \rightarrow 1$.
We also wish the impose a domain transformation, $\kappa(\jf)$, such that $0 \leq \kappa \leq 1$ and the individual \qnm{} frequencies are morphologically simplified.
\par For the domain transformation, inspection of \qnm{} frequencies with $\ell \leq 5$ suggest that
  \kappa(\jf) = \left( \ln( 2 - \jf ) / \ln(3) \right)^{1/(2-\ell+|m|)}
appropriately linearizes the sharp behavior of each frequency near $\jf=1$.
\par For the zero-damping constraint, when considering a \qnm{} frequency $\cw_{\lmn}$, zero-damping at $\jf=1$ implies that $\cw_{\lmn}(\kappa \approx 0)-m/2 \propto \kappa$, where $m/2$ is the well known limiting value for each \qnm{} frequencies real part as $\jf\rightarrow 1$.
% Plots of QNM frequencies

@@

  \begin{tabular}{lcr}
    \includegraphics[width=\figfactor\textwidth]{fig/fits_w.pdf} & \includegraphics[width=\figfactor\textwidth]{fig/fits_tau.pdf}
  \end{tabular}

@@

	\caption{ Fits of dimensionless \qnm{} central frequencies (solid lines) along with select numerical values (grey markers) computed using Leaver's method \cite{Leaver85}.
  Before the application of $\kappa(\j{})$, points are spaced between -\CwFitCalibrationRegion and \CwFitCalibrationRegion according to \CwFitCalibrationRegion times the $\sin$ of a fiducial angle which is uniformly spaced between $-\pi/2$ and $\pi/2$. Values of $\j{}$ are shown in the upper axis for $\kappa$ at $l=m$.
  The grey dashed line marks the value of $\kappa$ where $\j{}=0$. Fits of dimensionless \qnm{} decay rates (solid lines) along with select numerical values (grey markers) computed using Leaver's method \cite{Leaver85}. }
  \label{fig:qnms}

@@

% Plots of QNM frequencies
  \begin{tabular}{lcr}
    \includegraphics[width=\figfactor\textwidth]{fig/res_w.pdf} & \includegraphics[width=\figfactor\textwidth]{fig/res_tau.pdf}
  \end{tabular}
	\caption{ Fractional residual errors for fits of dimensionless \qnm{} central frequencies and decay times (solid lines) along with select numerical values (grey markers) computed using Leaver's method \cite{Leaver85}. }
  \label{fig:qnm_err}
  \label{eq:zd}
  \cw_{\lmn} = m/2 + \kappa \sum_{j=1}^{J} c_j \kappa^k \; .
In the case of the non-zero damped \qnm{s}, (e.g. $(\ell,m)=(3,2)$), a more general polynomial form may be adopted, namely,
  \label{eq:nzd}
  \cw_{\lmn} = \sum_{j=0}^{J} c_j \kappa^k \; .
\par The polynomial content of \eqn{eq:zd} and \eqn{eq:nzd} is determined by \gmvp{}.
\eqns{eq:cw_fit_1}{eq:cw_fit_9} display the resulting polynomial models.
Although each model's fractional error is within 2\%, each is error is dominated by well known nonlinear oscillations which are considered to be outside of the experimental accuracy of current \gw{} efforts.
\par \fig{fig:qnms} displays select training points, as well as model fits for $\cw_{lmn}$'s real and imaginary parts.
For the right and left panels, simple polynomial behavior of each curve is a result of the displayed linear domain in $\kappa(\jf)$.
In the left panel, we have scaled $1/\tau_{\lmn}$ by factors for $m/(n+1/2)$ to place the \qnm{s} with $n=0$ and $n=1$ at approximately the same scale.
% Equations for QNM frequency fits
\begin{widetext}
	\input{src/cwfit_eqns}
\end{widetext}
\subsection{Modeling spherical-spheroidal inner-products using \gmvr{}}
% Equations for mixing coefficients
\begin{widetext}
  \input{src/ysprod_eqns}
\end{widetext}

>>>>

\title{On modeling for Kerr black holes: Basis learning, QNM frequencies, and spherical-spheroidal mixing coefficients }

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Abstract %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
	%
	The ongoing direct detection of gravitational wave signals is aided by representative models of theoretical predictions. In particular, the process of model based detection, and subsequent comparison of signals the general relativity's predictions are aided by the modeling of information related to perturbed Kerr black holes. Here, we summarize recent methods and models for the analytically understood gravitational wave spectra (quasi-normal mode frequencies), and harmonic structure of Kerr black holes (mixing coefficients between spherical and spheroidal harmonics). Towards the construction of these models, two algorithms, GMVP and GMVR, for the automated polynomial and rational modeling of general dimensional complex scalars are presented. GMVR is a quasi-linear numerical method for interpolating rational functions. It therefore represents a solution to the problem of Runge's phenomenon.
	%
\end{abstract}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

From the first to second line of \eqn{eq:mvt}, we have used the definition of the exponential function (i.e. its series expansion).

@@

These adjustments of perspective result in

@@

\par The key to robustly solving for $\vec{\alpha}$ lies in the iterative refinement.

@@

\par We briefly review the application of \gmvr{} to a toy problem wherein a scalar rational function of two variables is treated.

@@

We then present two applications to \gw{s}.

@@

First we apply \gmvp{} to the modeling of complex valued Kerr \qnm{} frequencies.
Second, we apply \gmvr{} to the modeling of spin -2 spherical-spheroidal harmonic mixing coefficients.
While only 1D and 2D domains are treated here, we note that Ref. \cite{London:2018gaq} has used a version of \gmvp{} to model the \qnm{} excitation amplitudes in a 4D parameter space.

@@

Here we demonstrate \gmvr{}

@@

@@

  \includegraphics[width=\textwidth]{fig/issue3_summary.pdf}

@@

	\caption{ Standard summary plot for \gmvr. }

@@

%
\begin{figure}[htb]
  %
  \includegraphics[width=0.45\textwidth]{fig/issue3_pgreedy.pdf}
  %
	\caption{ Convergence of greedy process for \gmvr{} toy problem. }
  %
\end{figure}

%
\subsection{Modeling \qnm{} frequencies using \gmvp{}}
%
In seeking to apply \gmvp{} to select \qnm{} frequencies, we wish to impose a zero-damping constraint, namely that some frequencies are real as $\jf \rightarrow 1$.
%
We also wish the impose a domain transformation, $\kappa(\jf)$, such that $0 \leq \kappa \leq 1$ and the individual \qnm{} frequencies are morphologically simplified.
%
\par For the domain transformation, inspection of \qnm{} frequencies with $\ell \leq 5$ suggest that
%
\begin{align}
  \kappa(\jf) = \left( \ln( 2 - \jf ) / \ln(3) \right)^{1/(2-\ell+|m|)}
\end{align}
%
appropriately linearizes the sharp behavior of each frequency near $\jf=1$.
%
\par For the zero-damping constraint, when considering a \qnm{} frequency $\cw_{\lmn}$, zero-damping at $\jf=1$ implies that $\cw_{\lmn}(\kappa \approx 0)-m/2 \propto \kappa$, where $m/2$ is the well known limiting value for each \qnm{} frequencies real part as $\jf\rightarrow 1$.
%
% Plots of QNM frequencies
\begin{figure*}
  %
  \begin{tabular}{lcr}
    \includegraphics[width=\figfactor\textwidth]{fig/fits_w.pdf} & \includegraphics[width=\figfactor\textwidth]{fig/fits_tau.pdf}
  \end{tabular}
  %
	\caption{ Fits of dimensionless \qnm{} central frequencies (solid lines) along with select numerical values (grey markers) computed using Leaver's method \cite{Leaver85}.
  %
  Before the application of $\kappa(\j{})$, points are spaced between -\CwFitCalibrationRegion and \CwFitCalibrationRegion according to \CwFitCalibrationRegion times the $\sin$ of a fiducial angle which is uniformly spaced between $-\pi/2$ and $\pi/2$. Values of $\j{}$ are shown in the upper axis for $\kappa$ at $l=m$.
  %
  The grey dashed line marks the value of $\kappa$ where $\j{}=0$. Fits of dimensionless \qnm{} decay rates (solid lines) along with select numerical values (grey markers) computed using Leaver's method \cite{Leaver85}. }
  %
  \label{fig:qnms}
  %
\end{figure*}
%
% Plots of QNM frequencies
\begin{figure*}
  %
  \begin{tabular}{lcr}
    \includegraphics[width=\figfactor\textwidth]{fig/res_w.pdf} & \includegraphics[width=\figfactor\textwidth]{fig/res_tau.pdf}
  \end{tabular}
  %
	\caption{ Fractional residual errors for fits of dimensionless \qnm{} central frequencies and decay times (solid lines) along with select numerical values (grey markers) computed using Leaver's method \cite{Leaver85}. }
  %
  \label{fig:qnm_err}
  %
\end{figure*}
%
This implies that
%
\begin{align}
  \label{eq:zd}
  \cw_{\lmn} = m/2 + \kappa \sum_{j=1}^{J} c_j \kappa^k \; .
\end{align}
%
In the case of the non-zero damped \qnm{s}, (e.g. $(\ell,m)=(3,2)$), a more general polynomial form may be adopted, namely,
%
\begin{align}
  \label{eq:nzd}
  \cw_{\lmn} = \sum_{j=0}^{J} c_j \kappa^k \; .
\end{align}
%
\par The polynomial content of \eqn{eq:zd} and \eqn{eq:nzd} is determined by \gmvp{}.
%
\eqns{eq:cw_fit_1}{eq:cw_fit_9} display the resulting polynomial models.
%
Although each model's fractional error is within 2\%, each is error is dominated by well known nonlinear oscillations which are considered to be outside of the experimental accuracy of current \gw{} efforts.
%
\par \fig{fig:qnms} displays select training points, as well as model fits for $\cw_{lmn}$'s real and imaginary parts.
%
For the right and left panels, simple polynomial behavior of each curve is a result of the displayed linear domain in $\kappa(\jf)$.
%
In the left panel, we have scaled $1/\tau_{\lmn}$ by factors for $m/(n+1/2)$ to place the \qnm{s} with $n=0$ and $n=1$ at approximately the same scale.
%
% Equations for QNM frequency fits
\begin{widetext}
	\input{src/cwfit_eqns}
\end{widetext}
%
%
\subsection{Modeling spherical-spheroidal inner-products using \gmvr{}}
%
%
% Equations for mixing coefficients
\begin{widetext}
  \input{src/ysprod_eqns}
\end{widetext}
